---
title: "Homework 1 Pokemon"
author: "Emili Robles"
date: "2024-10-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Pokemon

```{r}
library(readr)
library(dplyr)
library(ggplot2)
pokemon <- read_csv("https://uwmadison.box.com/shared/static/hf5cmx3ew3ch0v6t0c2x56838er1lt2c.csv")

#a.
pokemon <- pokemon |>
  mutate(attack_to_defense = Attack / Defense)

#b.
median_ratios <- pokemon |>
  group_by(type_1) |>
  summarize(median_attack_to_defense = median(attack_to_defense, na.rm = TRUE)) |>
  arrange(desc(median_attack_to_defense))

#c.
pokemon <- pokemon |>
  mutate(type_1 = factor(type_1, levels = median_ratios$type_1))

attack_vs_defense <- ggplot(pokemon, aes(Attack, Defense)) +
  geom_point(alpha = 0.7) +
  facet_wrap(~type_1, scales = "free", ncol = 4) +
  labs(title = "Attack VS. Defense by Pokemon Types") + 
  theme_minimal()
attack_vs_defense

```

### d. A dynamic query-based visualization for this Pokemon dataset could be an interactive dashboard where users are able to explore different columns of characteristics of Pokemon other than just attack and defense. Questions this could answer could like, "How do different Pokemon types vary in their attack and defense statistics? The structure of this interaction could be having a dropdown menu to filter out the types of Pokemon, or also a filter menu to sort by Attack, Defense, the ratio, etc. The way we could display this for dynamic updates would be by providing a hover interaction to the users to be able to identify certain detailed stats that reveal a tooltip with the name, type, attack, etc. 