---
title: "Homework 1"
output: 
  html_document:
    self_contained: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T,eval=T,message=F,warning=F,fig.align='center')
library(tidyverse)
```


## Problem 1 <small>(2pts each, 8pts total)</small>

Here are a few probability exercises to get you warmed up.

a. Consider $X\sim \text{Normal}(5, 2^2)$. Find $P(4 < X < 7)$
```{r question-1a}
pnorm(7, mean=5, sd=2) - pnorm(4, mean = 5, sd =2)

```
b. Suppose you have an uneven 6-sided die where the even numbers are equally likely to occur and the odd numbers are also equally likely. The expected value of the die is $3.3333$ (i.e. $\frac{10}{3})$. What is the probability of getting a 6?
   (Hint: Let $p$ represent the probability of getting an even number, and represent the probability of the other numbers algebraically. Then derive an expression for the expected value, let it equal $\frac{10}{3}$, and solve. You can go back and check your work by plugging the result back into your original expressions and verify it works with R.)
```{r question-1b}
p <- 1/3
even_prob <- p/3
odd_prob <- (1-p)/3
ev <- 1*odd_prob + 2*even_prob + 3*odd_prob + 4*even_prob + 5*odd_prob + 6*even_prob
print(ev)
```
c. Give **two examples** of pairs of events that are mutually exclusive and explain why for each.
```{r question-1c}
#need to generate random outputs
#set.seed(100)
#make a coin flip 
flip <- sample(c('heads', 'tails'), size = 1, replace = TRUE)
if(flip == 'heads'){
  print("heads only printed out not tails, making it mutually exclusive.")
}else{
  print("tails only printed out, making it mutually exclusive") }
```
A coin flip with heads and tails is mutually exclusive because you can't get both as an outcome in one try, I proved it by making a simulation of it and it never pops up as both heads and tails, making it mutually exclusive.
```{r}
roll <- sample(c(1:6), size = 1, replace = TRUE)
roll
```
A dice is also mutually exclusive because you can get multiple numbers at the same time as shown by the code, only one number is able to pop up each time.

d. Give **two examples** of pairs of events that are independent and explain why for each.
```{r question-1d}
set.seed(100)
marbles <- sample(c('blue', 'black'), size = 2, replace= TRUE)
marbles
```
This is an independent set because I'm able to pull out marbles 2 times in a row and the outcome of the first draw doesn't affect the outcome of the second draw.
```{r}
set.seed(110)
lights <- sample(c('on','off') ,size = 2, replace = TRUE)
lights
```
This is an independent set because I have 2 lights in a room and I can turn both either on or off because the outcome from the first light doesn't affect whether to turn off or on to the other one.

## Problem 2 <small>(2pts each, 14pts total)</small>

For each of the following scenarios, decide what random variable may be appropriate to use as a model, and explain your reasoning as well as any assumptions you make.

a. The number of patients in an experimental drug trial that respond to treatment.
```{r question-2a}
#I'd use a binomial random variable because we have a certain number of patients in the experiment and we can assume that mostly there will be a percentage of them that respond to the treatment.
```


b. The number of red cars you observe driving through an intersection between 10am and 11am.
```{r question-2b}
#Poisson random variable because we essentially want to model how many red cars we see and because we are assuming that we will see some + we're given an exact timeframe of 1 hour, it makes it poisson.
```

c. What the second hand reads on the clock when you wake up from a long nap (approximately).
```{r question-2c}
#A uniform random variable because we're assuming that any second from 0 - 59 is equally likely so a uniform distribution would be able to get a model of the randomness of what time you wake up to.
```


d. How many people you need to swipe right on Tinder before you get a match.
```{r question-2d}
#Geometric random variable because assuming that it takes countless repeated rejections, we consider that number right before it lands on a match.
```

e. The length of time between mosquito bites a camper experiences while on a hike.
```{r question-2e}
#Exponential random variable, because this problem is like a poisson but assuming that we don't know the time that the hike will happen in, we have to use exponential instead of poisson.
```
Exponential random variable because because we dont know the exact time frame. 

f. Whether the Eagles win the Superbowl this year.
```{r question-2f}
#Bernoulli random variable because assuming that the Eagles are playing, we assume only 2 outcomes which is win or not win, which connects to a yes or no outcome.
```

g. The GPA of a randomly selected UW-Madison graduate.
```{r question-2g}
#Normal random variable because we're assuming that the GPA distributions are symmetric so we can do a normal model when we deal with the gpa averages.
```

## Problem 3 <small>(2pts each, 8pts total)</small>

Below you will find the plot of a probability density function for a random variable $X$.

```{r, echo=FALSE}
plot(x=c(0,1,2,3), y=c(0,.5,.5,0), type="l", ylab="density", xlab="x", main="pdf of X")
```

Using just geometry answer a and b:

a. What is $Pr[X < 1]$?
```{r question-3a}
area = (1/2)*1*.5
area
```

b. What is $Pr[1.5 < X < 2.5]$
```{r question-3b}
area1 = 1*0.3
area2 = 0.5*0.2
area3 = (1/2)*0.5*0.2
total = area1 + area2 + area3
total

```

c. Consider the CDF given below for random variable $Y$ (we're not naming it $X$ to keep it distinct from parts a and b)
```{r}
plot(x=0:6, y=c(0,.3,.5, .6, .75,.8,1), type="s", ylab="F(y)", xlab="y", main="cdf of Y")
abline(h=seq(.1,.9,.1), lty=2, col="gray")

```
Express the pmf of $Y$ in the following R code (in other words, fill in the missing probabilities)
```{r}
y<- 1:6
py <- c(.3,.2,.1,.15,.05,.2) ## edit this line
barplot(height=py, names=y, main="pmf of Y")
```
d. Calculate $E[Y]$ for the variable $Y$ in part c.
```{r question-3d}
sum = 1*.3 + 2*.2 + 3*.1+ 4*.15 +5*.05+ 6*.2
sum
```
